---
title: "Prediction Assignment"
author: "Alvina"
date: "Friday, July 24, 2015"
output: html_document
---

<b>PREPARING THE DATA:</b><br/>
Set the working directory.
```{r} 
setwd("F:/Ph.D/coursera") 
```
The dataset is downloaded from coursera website and saved to the working directory.Load the caret library.<br/>
```{r} 
library(caret)
```
Read the training and test dataset from the working directory.<br/>
```{r} 
training<-read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!"))
testing<-read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!"))
```
<b>EXPLORING THE DATA:</b><br/>
The number of observations and attributes in the training set is obtained.
```{r} 
dim(training)
```
The target variable is classe.
```{r} 
table(training$classe)
```
The number of classes in the dataset is noted.There are 5 classes and in that class A  is more in number.Among the 160 attributes there exists attributes which contains NA values and those which are not directly related to the target attribute.So those attributes are removed.<br/>
```{r} 
NA_Count = sapply(1:dim(training)[2],function(x)sum(is.na(training[,x])))
NA_list = which(NA_Count>0)
colnames(training[,c(1:7)])
```
After removing the above attributes the training dataset is again processed.<br/>
```{r} 
training = training[,-NA_list]
training = training[,-c(1:7)]
training$classe = factor(training$classe)
```
Similar processing is performed for testing dataset also.<b/>
```{r} 
testing = testing[,-NA_list]
testing = testing[,-c(1:7)]
``` 
<b>CROSS VALIDATION:</b><br/>
The problem is a classification problem.I have implemented random forest and regression method using 'rf' and 'rpart' in the caret package.The cross validation which is carried out is a 3-fold cross validation.The seed has been set to 1234.<br/>
The random forest model is generated.<br/>

```{r} 
set.seed(1234)
cv3 = trainControl(method="cv",number=3,allowParallel=TRUE,verboseIter=TRUE)
modrf = train(classe~., data=training, method="rf",trControl=cv3)
``` 
The regression tree model is generated.<b/>
```{r} 
modtree = train(classe~.,data=training,method="rpart",trControl=cv3)
``` 
Checking the performance.Table for random forest model.</b>
```{r} 
prf=predict(modrf,training)
ptree=predict(modtree,training)
acc <- table(prf,training$classe)
``` 
Accuracy is calculated.<br/>
```{r} 
confusionMatrix(acc)
``` 
Table for regression tree model.<br/>
```{r} 
table(ptree,training$classe)
``` 
Checking the performance for testing data.<br/>
```{r} 
prf=predict(modrf,testing)
ptree=predict(modtree,testing)
table(prf,ptree)
table(prf)
``` 
The results shows good accuracy for testing dataset with random forest model.From the above analysis I understood that random forest model is able to give better accuracy than regression tree model.<br/>
<b>SUBMISSION:</b><br/>
The random forest model is applied to testing dataset.<br/>
```{r} 
answers=predict(modrf,testing)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
answers
``` 
The predicted class for 20 test data are:
B A B A A E D B A A B C B A E E A B B B
